API design for machine learning software: experiences from the scikit-learn project

### 2013-06-27 paper.tex
s/choices/combinations/ at @mblondel's request

\begin{description}
  \item[Consistency.]
       All objects (basic or composite) share a consistent interface composed of
       a limited set of methods. This interface is documented in a consistent
       manner for all objects.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored and exposed as public attributes.
  \item[Non-proliferation of classes.]
       Learning algorithms are the only objects to be of custom classes.
       Datasets are represented as standard Python list, NumPy array or
       Scipy sparse matrices.
       Hyper-parameter names and values are represented as standard
       Python string or number whenever possible.
       This keeps \sklearn easy to use and easy to combine with other libraries.
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parametrized on other algorithms.
       Whenever feasible, such algorithms are implemented and composed from
       existing building blocks.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       an appropriate default value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

In most machine learning tasks, data is modeled as a set of variables.  For
example, in a classification or regression task, the goal is to find a mapping
between input variables $X_1, ..., X_p$, called features, and
some output variable $Y$. A sample is then defined as a pair of input/output
values $([x_1,..., x_p]^\mathrm{T}, y)$ of these variables. A widely used
representation of a dataset, a collection of
such samples, is a pair of matrix with numerical values: one for the input
values and one for the output values. Each row of these matrix
corresponds to one sample of the dataset and each column to one variable of
the problem.

@@

As introduced in Section~\ref{sec:estimators}, hyper-parameters set in the
constructor of an estimator
determine the behavior of the learning algorithm
and hence the performance of the resulting model on unseen data.
The problem of \textit{model selection} is therefore to find, within
some hyper-parameter space, the best combination of hyper-parameters, with
respect to some user-specified criterion. For example, a decision
tree with too small values of the maximal tree depth
parameter  will lead to underfitting, while too large
values will make it overfit.

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as input an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
set of hyperparameter settings to search through.
This set is represented as a mapping of parameter names
to a set of discrete choices in the case of grid search,
which exhaustively enumerates the ``grid'' (cartesian product)
of complete parameter choices.
Randomized search is a smarter algorithm
that avoids the combinatorial explosion in grid search
by sampling a fixed number of times from its parameter distributions
(see \citealt{bergstra2012}).

>>>>

\begin{description}
  \item[Consistency.]
       All objects (basic or composite) share a consistent interface composed of
       a limited set of methods. This interface is documented in a consistent
       manner for all objects.
  \item[Inspection.]
       Constructor parameters and parameter values determined by learning
       algorithms are stored and exposed as public attributes.
  \item[Non-proliferation of classes.]
       Learning algorithms are the only objects to be of custom classes.
       Datasets and hyper-parameter names and values are represented as standard
       Python, NumPy or SciPy objects whenever possible.
       This keeps \sklearn easy to use and easy to combine with other libraries.
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parametrized on other algorithms.
       Whenever feasible, such algorithms are implemented and composed from
       existing building blocks.
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       an appropriate default value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

@@

In most machine learning tasks, data is modeled as a set of variables.  For
example, in a classification or regression task, the goal is to find a mapping
between input variables $X_1, ..., X_p$, called features, and
some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
such samples.  As such, matrices of numerical values constitute a natural and
widely used representation in machine learning: each row in the
matrix corresponds to one sample of the dataset and each column to one
variable of the problem.

@@

As introduced in Section~\ref{sec:estimators}, hyper-parameters set in the
constructor of an estimator
determine the behavior of the learning algorithm
and hence the performance of the resulting model on unseen data.
The problem of \textit{model selection} is therefore to find, within
some hyper-parameter space, the best combination of hyper-parameters, with
respect to some user-specified criterion. For example, in the case of a decision
tree, too small values of the \texttt{max\_depth} parameter will make the tree
underfit, while too large values will make it overfit.

In \sklearn, model selection is supported in two distinct meta-estimators,
\texttt{GridSearchCV} and \texttt{RandomizedSearchCV}.  They take as input an
estimator (basic or composite), whose hyper-parameters must be optimized, and a
set of hyperparameter settings to search through.
This set is represented as a mapping of parameter names
to a set of discrete choices in the case of grid search,
which exhaustively enumerates the ``grid'' (cartesian product)
of complete parameter combinations.
Randomized search is a smarter algorithm
that avoids the combinatorial explosion in grid search
by sampling a fixed number of times from its parameter distributions
(see \citealt{bergstra2012}).
